---
title: "CSN 04 - Community Detection Algorithms"
author: "Carolina Jorge, Laura Cebollero"
date: "14th of November, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("task_01.R")
options(scipen=999)
```
```{r setup2, include=FALSE, echo=FALSE, cache=TRUE}
walktrapCommsWiki = computeWalktrap(wikiG)
```
## Introduction

In this lab we are asked to bla....


# 1. Metrics
We have ..........

```{r  echo=FALSE}
kable(karateMetricsTable)

```

## Discussion on metrics

# 2. Wikipedia community detection

Now we are going to detect the communities on the Wikipedia dataset provided in `wikipedia.gml`.

This whole network has X vertices and Y edges. Thus, we can see that this network is really large.

Because of this, Edge betweeness as a community detection algorithm has been immediately discarded
by us for its high computational demanding due to its inner working methods of removing one edge, computing the degrees, then restoring the removed edge and removing another one, compute again... And so on.

Looking at the table 1 in the Metrics section, we can see that the fastest community detection algorithms are walktrap, multilevel, label, infomap and fastgreedy.

Multilevel cannot be used because it needs the graph to be undirected and we are dealing, in fact,  with a directed graph.

We have chosen to perform the community detection with Walktrap because it works like []

Let's take a look at the total number of communities found with a function 
we have implemented:
```{r}
printNumberOfCommunitiesFound(walktrapCommsWiki)
```

We have tried to plot the whole graph but of course nothing can really be determined
by looking at it, because it is very dense.

TODO: Put graphic

Because of this we are only going to take a look at the first 3 communities detected.
```{r   out.width="350px", out.height="350px", fig.align='center'}
x = 3
plotGraphFirstXCommunities(walktrapCommsWiki, wikiG, x)
```

# Summary

TODO: write discussions of problems encountered & things learned